---
title: 'POLS 207 - Problem Set #2'
author: "Cori Lopazanski"
date: "4/15/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
library(tidyverse)

```

### Problem 1

```{r}
### Read Olken Data & Initial Cleaning
olken <- read_csv("olken_data.csv") %>% 
  drop_na() # Remove all observations with NA

```

a) Use difference in means estimator to estimate ATE of community participation 

```{r}
# Calculate average percent missing for each group (treatment and control):
means <- olken %>% 
  group_by(treat_invite) %>% 
  summarize(mean = mean(pct_missing)) 

# Find difference in means:
ate <- means$mean[means$treat_invite == 1] - means$mean[means$treat_invite == 0]
```

Encouraging community participation in monitoring had an average treatment effect of `r ate` on the percent expenditures missing for road building projects. 

b) what characteristics do the treatment/control groups differ?

c) use regression to estimate the ATE (heteroskedastistic-consistent SEs) is this different from DIMs estimate?
d)
e)

### Problem 2

a)
b)
c)

### Problem 3

a)
b)